---
title: "BIOST 544A Course Project: WHO Life Expectancy"
author: "My-Anh Doan and Katie Denecke"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

| 1. Introduction
|    1.1 Variable description
|    1.2 Analysis questions and aims
|    1.3 Analysis approach
| 2. Data preparation
| 3. Initial exploration
|    3.1 Missing data
|    3.2 Test/train split
| 4. Main data analysis
|    4.1 Simple linear regression models
|        4.1.1 Alcohol
|        4.1.2 HepB Immunization
|        4.1.3 Polio Immunization
|        4.1.4 Diphtheria Immunization
|        4.1.5 HIV/AIDS Deaths
|    4.2 Multiple linear regression models
|        4.2.1 Model fitting
|        4.2.2 Model selection
| 5. Final results and interpretations

# 1. Introduction
This publicly available dataset originates from the Global Health Observatory (GHO) data repository under the World Health Organization (WHO). The dataset contains data on life expectancy, health factors, and economic data (collected by the United Nation website) from 2000 to 2015 for 193 countries. There is a total of 22 attributes in this dataset that can be divided into broad categories related to immunization, mortality, economical, and social factors. The dataset aims to determine significant factors that affect life expectancy.

## 1.1 Variable description
For this course project, we will only look at five predictor variables in relation to life expectancy: alcohol consumption, HepB immunization, polio immunization, diphtheria immunization, and HIV/AIDS death.

Factor | Variable | Type | Description
------ | -------- | -----| ----------- 
Life expectancy | Response | Quantitative | Life expectancy in years
Alcohol | Predictor | Quantitative | Alcohol consumption (liters of pure alcohol; per capita)
Hepatitis B | Predictor | Quantitative | Hepatitis B (HepB) immunization coverage among 1-year-olds (%)
Polio | Predictor | Quantitative | Polio immunization coverage among 1-year-olds (%)
Diphtheria | Predictor | Quantitative | TDAP immunization coverage among 1-year-olds (%)
HIV/AIDS | Predictor | Quantitative | Deaths per 1000 live births (0-4 years) caused by HIV/AIDS

## 1.2 Analysis questions and aims
The dataset aims to answer the following questions: 

1. Among the five predictors chosen above, which ones are actually significant to life expectancy?
2. Is there a positive or negative relationship between life expectancy and alcohol consumption?
3. What is the impact of immunization coverage on life expectancy? Which immunization has a significant impact on life expectancy?
4. How does HIV/AIDS mortality rate affect life expectancy?

## 1.3 General analysis approach
We will use regression tools and techniques to determine the significant predictors of life expectancy. 

# 2. Data preparation
```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(janitor) # Data cleaning package

# Load data
data <- read.csv("./data/Life Expectancy Data.csv")

data <- data %>%
  clean_names() %>% 
  select(life_expectancy, alcohol, hepatitis_b, polio, diphtheria, hiv_aids) %>% 
  mutate_all(as.numeric)

str(data)
```

# 3. Initial exploration

## 3.1 Handling missing data in predictor variables
```{r message = FALSE, warning = FALSE}
library(naniar)
library(missForest)

# Check for any observations/rows with missing data
sum(complete.cases(data) == FALSE) / nrow(data)    # proportion of missing data
missing_vals <- which(complete.cases(data) == FALSE)    # rows with missing data

# Visualization of missing values
gg_miss_var(data, show_pct = TRUE)

# Impute missing values in covariates
set.seed(1001)
data_imputed_covars <- missForest(xmis = subset(data, select = -c(life_expectancy)),
                                  maxiter = 50)$ximp

data_imputed <- cbind(data$life_expectancy, data_imputed_covars)
colnames(data_imputed)[1] = "life_expectancy"
```

## 3.2 Train-test split dataset
``` {r, message = FALSE, warning = FALSE}
set.seed(1)
rand_sample <- sample(1:nrow(data_imputed),
                      size = 0.7 * nrow(data_imputed),
                      replace = FALSE)

training_data <- data_imputed[rand_sample, ]
test_data <- data_imputed[-rand_sample, ]
```

# 4. Main data analysis

## 4.1 Simple linear regression models
First, we look at the individual predictors and their effect on life expectancy.

### 4.1.1 Alcohol
```{r message = FALSE, warning = FALSE}
plot(x = training_data$alcohol, y = training_data$life_expectancy)

# Fit model to training data
alc_model_lrm <- glm(as.factor(life_expectancy) ~ alcohol, family = "binomial",
                     data=training_data)  
summary(alc_model_lrm)

# Use fitted model on test data
model_lrm <- glm(as.factor(life_expectancy) ~ alcohol,  family = "binomial",
                 data=test_data)
summary(model_lrm)
```

### 4.1.2 HepB Immunization
```{r message = FALSE, warning = FALSE}
plot(x = training_data$hepatitis_b, y = training_data$life_expectancy)

# Fit model to training data
hepB_model_lrm <- glm(as.factor(life_expectancy) ~ hepatitis_b, family = "binomial",
                 data=training_data)  
summary(hepB_model_lrm)

# Use fitted model on test data
model_lrm <- glm(as.factor(life_expectancy) ~ hepatitis_b,  family = "binomial",
                 data=test_data)
summary(model_lrm)
```

### 4.1.3 Polio Immunization
```{r message = FALSE, warning = FALSE}
plot(x = training_data$polio, y = training_data$life_expectancy)

# Fit model to training data
polio_model_lrm <- glm(as.factor(life_expectancy) ~ polio, family = "binomial",
                 data=training_data)  
summary(polio_model_lrm)

# Use fitted model on test data
model_lrm <- glm(as.factor(life_expectancy) ~ polio,  family = "binomial",
                 data=test_data)
summary(model_lrm)
```

### 4.1.4 Diphtheria Immunization
```{r message = FALSE, warning = FALSE}
plot(x = training_data$diphtheria, y = training_data$life_expectancy)

# Fit model to training data
diph_model_lrm <- glm(as.factor(life_expectancy) ~ diphtheria, family = "binomial",
                 data=training_data)  
summary(diph_model_lrm)

# Use fitted model on test data
model_lrm <- glm(as.factor(life_expectancy) ~ diphtheria,  family = "binomial",
                 data=test_data)
summary(model_lrm)
```

### 4.1.5 HIV/AIDS Deaths
```{r message = FALSE, warning = FALSE}
plot(x = training_data$hiv_aids, y = training_data$life_expectancy)

# Fit model to training data
hiv_model_lrm <- glm(as.factor(life_expectancy) ~ hiv_aids, family = "binomial",
                 data=training_data)  
summary(hiv_model_lrm)

# Use fitted model on test data
model_lrm <- glm(as.factor(life_expectancy) ~ hiv_aids,  family = "binomial",
                 data=test_data)
summary(model_lrm)
```

## 4.2 Multiple linear regression models

### 4.2.1 Model fitting
```{r message = FALSE, warning = FALSE}
library(car)

# Simple linear regression model using all variables, fitted on training data
lm_all <- lm(life_expectancy ~ ., data = training_data)
summary(lm_all)
vif(lm_all) # no multicollinearity in our data

# Perform stepwise forward selection using AIC as selection criteria on training data
lm_intercept_only <- lm(life_expectancy ~ 1 , data = training_data)
summary(lm_intercept_only)
  
forward_AIC <- step(lm_intercept_only, scope = formula(lm_all),
                    direction = "forward", trace = 0, k = 2)
forward_AIC$anova

# Perform stepwise forward selection using BIC as selction criteria on training data
forward_BIC <- step(lm_intercept_only, scope = formula(lm_all),
                    direction = "forward", trace = 0, k = log(nrow(training_data)))
forward_BIC$anova

forward_AIC$coefficients
forward_BIC$coefficients

anova(forward_BIC, forward_AIC)

# Adding a polynomial on most significant predictor (based on SLR model p-values)
# Builds upon the stepwise forward selection model using AIC

M2 <- lm(life_expectancy ~ poly(hiv_aids, degree = 4, raw = TRUE) + diphtheria
         + alcohol + polio + hepatitis_b, data = training_data)

anova(forward_AIC, M2)
```

### 4.2.2 Model selection
```{r message = FALSE, warning = FALSE}
anova(forward_AIC, M2)

summary(forward_AIC)
summary(M2)

# Use fitted model on test data
test_pred <- predict(M2, newdata = test_data[ , -1])
test_obs_pred <- cbind(test_data$life_expectancy, test_pred)
colnames(test_obs_pred) <- c("Observed", "Predicted")

head(test_obs_pred)
```

# 5. Results and interpretations



